### 2018-06-15 paper/4-build.tex
Work on S5

\vspace{-2mm}
\subsubsection{Reordering}\label{sec-reordering}

@@

\vspace{-2mm}
\subsubsection{Recursive}\label{sec-recursive}

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological\hspace{2mm}\S\ref{sec-topological} & Reordering\hspace{2mm}\S\ref{sec-reordering} & Recursive\hspace{2mm}\S\ref{sec-recursive}    \\\hline
\hline
Dirty bit\hfill\S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake*  \\\hline
Verifying trace\hfill\S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel*            & \Shake                      \\\hline
Constructive trace\hspace{2mm}\hfill\S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel*             & Cloud \Shake*                \\\hline
\end{tabular}
\vspace{2mm}
\caption{Build systems \`a la carte. Systems marked * are hypothetical systems that do not currently exist.\label{tab-build-systems}}
\vspace{-2mm}
\end{table}

>>>>

\vspace{-2mm}
\subsubsection{Reordering}\label{sec-restarting}

@@

\vspace{-2mm}
\subsubsection{Recursive}\label{sec-suspending}

@@

\begin{table}[h]
\smaller
\centering
\begin{tabular}{l||c|c|c}
\hline
Property           & Topological\hspace{2mm}\S\ref{sec-topological} & Restarting\hspace{2mm}\S\ref{sec-restarting} & Suspending\hspace{2mm}\S\ref{sec-suspending}    \\\hline
\hline
Dirty bit\hfill\S\ref{sec-dirty-bit}                    & \Make    & \Excel                   & Approximate \Shake*  \\\hline
Verifying trace\hfill\S\ref{sec-verifying-traces}       & \Ninja   & Traced \Excel*            & \Shake                      \\\hline
Constructive trace\hspace{2mm}\hfill\S\ref{sec-constructive-traces} & \Bazel   & Cloud \Excel*             & Cloud \Shake*                \\\hline
\end{tabular}
\vspace{2mm}
\caption{Build systems \`a la carte. Systems marked * are hypothetical systems that do not currently exist.\label{tab-build-systems}}
\vspace{-2mm}
\end{table}

### 2018-06-15 paper/5-implementations.tex
Work on S5

In the previous sections we discussed the types of build systems, and how they
can be broken down. But these divisions were not obvious to us, and only by
concretely implementing and refactoring each build system did we determine the
underlying commonalities. In this section we share some of the code that got us
there.

@@

We provide an implementation of \Make using our framework in
Fig.~\ref{fig-make-implementation}. \Make processes keys in a linear order based
on a topological sort (see~\S\ref{sec-parallelism} for parallel \Make). For
each key, it builds it if it is older than any of its dependencies. We capture
the persistent build information that \Make stores by a pair
\hs{(@@modTime,}~\hs{now)} comprising the \emph{file modification time} function
\hs{modTime}~\hs{::}~\hs{k}~\hs{->}~\hs{Time} and the \emph{current time} \hs{now}.
Setting aside the explicit manipulation of file modification times, which in
reality is taken care of by the file system, the function \hs{make} captures
the essence of \Make in a clear and precise manner.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: current time and file modification times
type Time       = Integer
type MakeInfo k = (Time, Map k Time)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system
make :: Ord k => Build Applicative (MakeInfo k) k v
make = topological modTimeRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Topological scheduler
topological :: Ord k => Rebuilder Applicative i k v -> Build Applicative i k v
topological rebuilder tasks key = execState $ forM_ chain $ \k -> case tasks k of
    Nothing   -> return ()
    Just task -> do
        value <- gets (getValue k)
        let newTask :: Task (MonadState i) k v
            newTask = rebuilder k value task
            newFetch :: k -> StateT i (State (Store i k v)) v
            newFetch = lift . gets . getValue
        info <- gets getInfo
        (newValue, newInfo) <- runStateT (run newTask newFetch) info
        modify $ putInfo newInfo . putValue k newValue
  where
    deps k = case tasks k of { Nothing -> []; Just task -> dependencies task }
    chain  = case topSort (reachable deps key) of
        Nothing -> error "Cannot build tasks with cyclic dependencies"
        Just xs -> xs
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Ord k => (k -> [k]) -> k -> Graph k
topSort   :: Ord k => Graph k -> Maybe [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on file modification times
modTimeRebuilder :: Ord k => Rebuilder Applicative (MakeInfo k) k v
modTimeRebuilder key value task = Task $ \fetch -> do
    (now, modTime) <- get
    let dirty = case Map.lookup key modTime of
            Nothing -> True
            time -> any (\d -> Map.lookup d modTime > time) (dependencies task)
    if not dirty
    then return value
    else do
        put (now + 1, Map.insert key now modTime)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximate task dependencies: a map of known upper bounds of dependencies
type ApproximateDependencies k = Map k [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: dirty keys, approximate dependencies, and calculation chain
type CalcChain k = [k]
type ExcelInfo k = ((Set k, ApproximateDependencies k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = restarting approximateRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Restarting scheduler
restarting :: Ord k => Rebuilder Monad i k v -> Build Monad (i, Chain k) k v
restarting rebuilder tasks key = execState $ do
    chain    <- gets (snd . getInfo)
    newChain <- go Set.empty $ chain ++ [key | key `notElem` chain]
    modify . mapInfo $ \(i, _) -> (i, newChain)
  where
    go :: Set k -> Chain k -> State (Store (i, Chain k) k v) (Chain k)
    go _    []     = return []
    go done (k:ks) = case tasks k of
        Nothing -> (k :) <$> go (Set.insert k done) ks
        Just task -> do
            store <- get
            let value = getValue k store
                newTask :: Task (MonadState i) k (Either k v)
                newTask = try $ rebuilder k value task -- `try' is defined below
                newFetch :: k -> State i (Either k v)
                newFetch k | k `Set.member` done = return $ Right (getValue k store)
                           | otherwise           = return $ Left k
            case runState (run newTask newFetch) (fst $ getInfo store) of
                (Left dep, _) -> go done $ [ dep | dep `notElem` ks ] ++ ks ++ [k]
                (Right newValue, newInfo) -> do
                    modify $ putInfo (newInfo, []) . updateValue k value newValue
                    (k :) <$> go (Set.insert k done) ks
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a total task into a task accepting a partial fetch callback
try :: Task (MonadState i) k v -> Task (MonadState i) k (Either e v)
try task = Task $ \fetch -> runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on approximate dependencies
approximateRebuilder :: Ord k => Rebuilder Monad (Set k, ApproximateDependencies k) k v
approximateRebuilder key value task = Task $ \fetch -> do
    (isDirty, deps) <- get
    let dirty = key `Set.member` isDirty || case Map.lookup key deps of
                                                Nothing -> True
                                                Just ks -> any (`Set.member` isDirty) ks
    if not dirty
    then return value
    else do
        put (Set.insert key isDirty, deps)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = recursive vtRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive :: Ord k => Rebuilder Monad i k v -> Build Monad i k v
recursive rebuilder tasks key store = fst $ execState (fetch key) (store, Set.empty)
  where
    fetch :: k -> State (Store i k v, Set k) v
    fetch key = case tasks key of
        Nothing -> gets (getValue key . fst)
        Just task -> do
            done <- gets snd
            when (key `Set.notMember` done) $ do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                    newFetch :: k -> StateT i (State (Store i k v, Set k)) v
                    newFetch k = do v <- lift $ fetch k
                                    put =<< lift (gets (getInfo . fst))
                                    return v
                info <- gets (getInfo . fst)
                (newValue, newInfo) <- runStateT (run newTask newFetch) info
                modify $ \(s, _) ->
                    (putInfo newInfo $ putValue key newValue s, Set.insert key done)
            gets (getValue key . fst)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    vt <- get
    dirty <- not <$> verifyVT key value (fmap hash . fetch) vt
    if not dirty
    then return value
    else do
        (newValue, deps) <- trackM task fetch
        put =<< recordVT key newValue deps (fmap hash . fetch) =<< get
        return newValue
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{1mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Bazel build system, implementation of 'restarting2' is omitted (22 lines)
bazel :: (Ord k, Hashable v) => Build Monad (CT k v) k v
bazel = restarting2 ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Cloud Shake build system, implementation of 'recursive' is given in Fig. 8
cloudShake :: (Ord k, Hashable v) => Build Monad (CT k v) k v
cloudShake = recursive ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on constructive traces
ctRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (CT k v) k v
ctRebuilder key value task = Task $ \fetch -> do
    ct <- get
    maybeCachedValue <- constructCT key value (fmap hash . fetch) ct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- track task fetch
            put =<< recordCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Bazel and Cloud \Shake using our framework;
\hs{restarting2} differs from the restarting scheduler used by Excel
(Fig.~\ref{fig-excel-implementation}) in that it does not rely on the
calculation chain.}
% \hs{topological} is defined in Fig.~\ref{fig-make-implementation}.
\label{fig-bazel-cloud-shake-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Buck build system, implementation of 'topological' is given in Fig. 6
buck :: (Hashable k, Hashable v) => Build Applicative (DCT k v) k v
buck = topological (unliftRebuilder dctRebuilder)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a monadic rebuilder to the corresponding applicative one
unliftRebuilder :: Rebuilder Monad i k v -> Rebuilder Applicative i k v
unliftRebuilder rebuilder key value task = rebuilder key value $ Task $ run task
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Nix build system, implementation of 'recursive' is given in Fig. 8
nix :: (Hashable k, Hashable v) => Build Monad (DCT k v) k v
nix = recursive dctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on deterministic constructive traces
dctRebuilder :: (Hashable k, Hashable v) => Rebuilder Monad (DCT k v) k v
dctRebuilder key _value task = Task $ \fetch -> do
    dct <- get
    maybeCachedValue <- constructDCT key (fmap hash . fetch) dct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- trackM task fetch
            put =<< recordDCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Buck and \Nix using our framework.}
\label{fig-buck-nix-implementation}
\vspace{-2mm}
\end{figure}

>>>>

In the previous sections we discussed the types of build systems, and how they
can be broken down into two main components: a scheduler and a rebuilder. This
decomposition was not obvious to us, and only by concretely implementing and
refactoring each build system did we determine the underlying commonalities.
In this section we share some of the code that got us there.

@@

We provide an implementation of \Make using our framework in
Fig.~\ref{fig-make-implementation}. \Make processes keys in a linear order based
on a topological sort (see~\S\ref{sec-parallelism} for parallel \Make). For
each key, it builds it if it is older than any of its dependencies. We capture
the persistent build information that \Make stores by a pair
\hs{(@@now,}~\hs{modTime)} comprising the \emph{current time} \hs{now} and the
dictionary \hs{modTime}~\hs{::}~\hs{Map}~\hs{k}~\hs{Time} that stores \emph{file
modification times}. The top-level definition
\hs{make}~\hs{=}~\hs{topological}~\hs{modTimeRebuilder} is a direct translation
of the scheduler-rebuilder decomposition, illustrated in Table~\ref{tab-build-systems}.

Setting aside the explicit manipulation of file modification times, which in
reality is taken care of by the file system, the function \hs{make} captures
the essence of \Make in a clear and precise manner.

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: current time and file modification times
type Time       = Integer
type MakeInfo k = (Time, Map k Time)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Make build system
make :: Ord k => Build Applicative (MakeInfo k) k v
make = topological modTimeRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A topological task scheduler
topological :: Ord k => Rebuilder Applicative i k v -> Build Applicative i k v
topological rebuilder tasks key = execState $ forM_ chain $ \k -> case tasks k of
    Nothing   -> return ()
    Just task -> do
        value <- gets (getValue k)
        let newTask :: Task (MonadState i) k v
            newTask = rebuilder k value task
            newFetch :: k -> StateT i (State (Store i k v)) v
            newFetch = lift . gets . getValue
        info <- gets getInfo
        (newValue, newInfo) <- runStateT (run newTask newFetch) info
        modify $ putInfo newInfo . putValue k newValue
  where
    deps k = case tasks k of { Nothing -> []; Just task -> dependencies task }
    chain  = case topSort (reachable deps key) of
        Nothing -> error "Cannot build tasks with cyclic dependencies"
        Just xs -> xs
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Standard graph algorithms (implementation omitted)
reachable :: Ord k => (k -> [k]) -> k -> Graph k
topSort   :: Ord k => Graph k -> Maybe [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on file modification times
modTimeRebuilder :: Ord k => Rebuilder Applicative (MakeInfo k) k v
modTimeRebuilder key value task = Task $ \fetch -> do
    (now, modTime) <- get
    let dirty = case Map.lookup key modTime of
            Nothing -> True
            time -> any (\d -> Map.lookup d modTime > time) (dependencies task)
    if not dirty
    then return value
    else do
        put (now + 1, Map.insert key now modTime)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Make using our framework.}\label{fig-make-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximate task dependencies: a map of known upper bounds of dependencies
type ApproximateDependencies k = Map k [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: dirty keys, approximate dependencies, and calculation chain
type CalcChain k = [k]
type ExcelInfo k = ((Set k, ApproximateDependencies k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = restarting approximateRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A restarting task scheduler
restarting :: Ord k => Rebuilder Monad i k v -> Build Monad (i, Chain k) k v
restarting rebuilder tasks key = execState $ do
    chain    <- gets (snd . getInfo)
    newChain <- go Set.empty $ chain ++ [key | key `notElem` chain]
    modify . mapInfo $ \(i, _) -> (i, newChain)
  where
    go :: Set k -> Chain k -> State (Store (i, Chain k) k v) (Chain k)
    go _    []     = return []
    go done (k:ks) = case tasks k of
        Nothing -> (k :) <$> go (Set.insert k done) ks
        Just task -> do
            store <- get
            let value = getValue k store
                newTask :: Task (MonadState i) k (Either k v)
                newTask = try $ rebuilder k value task -- `try' is defined below
                newFetch :: k -> State i (Either k v)
                newFetch k | k `Set.member` done = return $ Right (getValue k store)
                           | otherwise           = return $ Left k
            case runState (run newTask newFetch) (fst $ getInfo store) of
                (Left dep, _) -> go done $ [ dep | dep `notElem` ks ] ++ ks ++ [k]
                (Right newValue, newInfo) -> do
                    modify $ putInfo (newInfo, []) . updateValue k value newValue
                    (k :) <$> go (Set.insert k done) ks
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a total task into a task accepting a partial fetch callback
try :: Task (MonadState i) k v -> Task (MonadState i) k (Either e v)
try task = Task $ \fetch -> runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A task rebuilder based on approximate dependencies
approximateRebuilder :: Ord k => Rebuilder Monad (Set k, ApproximateDependencies k) k v
approximateRebuilder key value task = Task $ \fetch -> do
    (isDirty, deps) <- get
    let dirty = key `Set.member` isDirty || case Map.lookup key deps of
                                                Nothing -> True
                                                Just ks -> any (`Set.member` isDirty) ks
    if not dirty
    then return value
    else do
        put (Set.insert key isDirty, deps)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = suspending vtRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- A suspending task scheduler
suspending :: Ord k => Rebuilder Monad i k v -> Build Monad i k v
suspending rebuilder tasks key store = fst $ execState (fetch key) (store, Set.empty)
  where
    fetch :: k -> State (Store i k v, Set k) v
    fetch key = case tasks key of
        Nothing -> gets (getValue key . fst)
        Just task -> do
            done <- gets snd
            when (key `Set.notMember` done) $ do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                    newFetch :: k -> StateT i (State (Store i k v, Set k)) v
                    newFetch k = do v <- lift $ fetch k
                                    put =<< lift (gets (getInfo . fst))
                                    return v
                info <- gets (getInfo . fst)
                (newValue, newInfo) <- runStateT (run newTask newFetch) info
                modify $ \(s, _) ->
                    (putInfo newInfo $ putValue key newValue s, Set.insert key done)
            gets (getValue key . fst)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    vt <- get
    dirty <- not <$> verifyVT key value (fmap hash . fetch) vt
    if not dirty
    then return value
    else do
        (newValue, deps) <- trackM task fetch
        put =<< recordVT key newValue deps (fmap hash . fetch) =<< get
        return newValue
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{1mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Bazel build system, implementation of 'restarting2' is omitted (22 lines)
bazel :: (Ord k, Hashable v) => Build Monad (CT k v) k v
bazel = restarting2 ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Cloud Shake build system, implementation of 'suspending' is given in Fig. 8
cloudShake :: (Ord k, Hashable v) => Build Monad (CT k v) k v
cloudShake = suspending ctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on constructive traces
ctRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (CT k v) k v
ctRebuilder key value task = Task $ \fetch -> do
    ct <- get
    maybeCachedValue <- constructCT key value (fmap hash . fetch) ct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- track task fetch
            put =<< recordCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Bazel and Cloud \Shake using our framework;
\hs{restarting2} differs from the restarting scheduler used by Excel
(Fig.~\ref{fig-excel-implementation}) in that it does not rely on the
calculation chain.}
% \hs{topological} is defined in Fig.~\ref{fig-make-implementation}.
\label{fig-bazel-cloud-shake-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Buck build system, implementation of 'topological' is given in Fig. 6
buck :: (Hashable k, Hashable v) => Build Applicative (DCT k v) k v
buck = topological (unliftRebuilder dctRebuilder)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a monadic rebuilder to the corresponding applicative one
unliftRebuilder :: Rebuilder Monad i k v -> Rebuilder Applicative i k v
unliftRebuilder rebuilder key value task = rebuilder key value $ Task $ run task
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Nix build system, implementation of 'suspending' is given in Fig. 8
nix :: (Hashable k, Hashable v) => Build Monad (DCT k v) k v
nix = suspending dctRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on deterministic constructive traces
dctRebuilder :: (Hashable k, Hashable v) => Rebuilder Monad (DCT k v) k v
dctRebuilder key _value task = Task $ \fetch -> do
    dct <- get
    maybeCachedValue <- constructDCT key (fmap hash . fetch) dct
    case maybeCachedValue of
        Just cachedValue -> return cachedValue
        Nothing -> do
            (newValue, deps) <- trackM task fetch
            put =<< recordDCT key newValue deps (fmap hash . fetch) =<< get
            return newValue
\end{minted}
\vspace{-2mm}
\caption{Implementations of \Buck and \Nix using our framework.}
\label{fig-buck-nix-implementation}
\vspace{-2mm}
\end{figure}
